Open source landscape ecology tools

### 2020-07-26 paper/paper.Rmd
adds new ideas
splits lines
improves spatial data processing section

While R has several build-in data structures, including data frames, it has not got any internal support for reading, processing, or visualizing spatial data.
However, there was a notable interest in spatial data analysis since the inception of this language, which is shown in a long R history of packages supporting spatial data (\cite{Lovelace2019}, page 10). <!--history section-->
Comparing to dedicated GIS software (e.g., QGIS, ArcGIS), both raster and vector models have several implementations in the R language.
<!-- are pros/cons of that needed? -->
<!-- info about GDAL/PROJ? -->

@@

<!-- elevatr + ?? - elevation -->
<!-- land cover -->
<!-- rgbif + ?? -->
<!-- rgee -->

@@

Finally, there is a number of tools for landscape ecology implemented in some GIS software, such as ... for GRASS GIS or ... for ... .
Gladly, it is possible to control several GIS software directly from R using dedicated packages, such as *rgrass7* for ..., ... for ..., and *rgee* for Google Earth Engine.

@@

>>>>

While R has several build-in data structures, including data frames, it has not got any internal support for reading, processing, or visualizing spatial data.
However, there was a notable interest in spatial data analysis since the inception of this language, which is shown in a long R history of packages supporting spatial data (\cite{Lovelace2019}, page 10).
Comparing to dedicated GIS software (e.g., QGIS, ArcGIS), both raster and vector models have several implementations in the R language.
<!-- are pros/cons of that needed? -->
<!-- info about GDAL/PROJ? -->

@@

<!-- elevatr + ?? - elevation -->
<!-- land cover -->
<!-- rgbif + ?? -->
<!-- climate data? -->
<!-- https://github.com/rspatial/geodata/blob/master/R/worldclim.R -->
<!-- rgee -->
<!-- look for rOpenSci packages -->

@@

Finally, there is a number of tools for landscape ecology implemented in GIS software, such as r.li or r.pi for GRASS GIS \cite{wegmann2018r,neteler2012grass}<!--ADD R.LI CITATION-->, terrain analysis methods in SAGA GIS \cite{gmd-8-1991-2015}, or morphological operations for Google Earth Engine.<!--https://github.com/r-spatial/rgee/blob/examples/image/morphological_operations.R-->
Gladly, it is possible to control several GIS software directly from R using dedicated packages, such as *rgrass7* \cite{R-grass7} for GRASS GIS, *RSAGA* \cite{R-RSAGA} for SAGA GIS, and *rgee* \cite{R-rgee} for Google Earth Engine.

<!--JN: should we mention ArcGIS here and some of its modules?-->
<!--JN: in the same vain as above: should we mention Google Earth Engine?-->

@@

<!-- missing packages: data sources, such as global land cover, worldclim 2.0, etc. -->
